Stream of Consciousness Literature: A 'Joyceless' Linguistic Landscape

### 2024-04-04 paper/paper.qmd
paper update

---
title: "Stream of Consciousness Literature: A 'Joyceless' Linguistic Landscape" 
subtitle: "Exploring Word Frequency, Sentiment Value and Mental Health Themes in the Works of Joyce, Woolf, Proust, Mansfield and Eliot from Project Gutenberg"
author: Quang Mai
thanks: "Code and data are available at: https://github.com/ponolite/stream-consciousness-language.git"
date: today
date-format: long
abstract: "This project focuses on understanding the language used by renowned stream of consciousness (SOC) authors James Joyce, Virginia Woolf, Marcel Proust, Katherine Mansfield and T.S Eliot. By conducting word frequency analysis and sentiment analysis of these authors' nine novels, this paper aims to uncover shared linguistic patterns and gain insights into the authors' mental states. Ultimatlely, this paper attempts to offer insights into themes of self-identity, anxiety, disassociation and existential contemplation within Western society and its literary circle from late 19th to mid-20th century. (add one sentence on main results)"
format: pdf
number-sections: true
bibliography: references.bib
toc: true
---

@@

@@

@@

@@

>>>>

---
title: "Stream of Consciousness Literature: A 'Joyceless' Linguistic Landscape" 
subtitle: "Exploring Word Frequency, Sentiment Value and Mental Health Themes in the Works of Joyce, Woolf, Proust, Mansfield and Eliot from Project Gutenberg"
author: Quang Mai
thanks: "Code and data are available at: https://github.com/ponolite/stream-consciousness-language.git"
date: today
date-format: long
abstract: "This project focuses on understanding the language used by renowned stream of consciousness (SOC) authors James Joyce, Virginia Woolf, Marcel Proust, Katherine Mansfield and T.S Eliot. By conducting word frequency analysis and sentiment analysis on these authors' nine novels, this paper aims to uncover shared linguistic patterns and gain insights into the authors' mental states. Ultimately, this paper attempts to offer insights into themes of self-identity, anxiety, disassociation and existential contemplation within Western society and its literary circle from late 19th to mid-20th century. (add one sentence on main results)"
format: pdf
number-sections: true
bibliography: references.bib
toc: true
---

@@

This includes words such as chapter titles, Roman numbers and personal names, affecting the integrity of data analysis ("Swann" being the most common word)

@@

Words are singled out and analyzed without context which could have affected their intended meanings, especially in such a complex genre such as stream of consciousness. The limiting novel editions also doesn't make sure that their literary integrity are maintained and the analysis might have missed important texts of other editions.

@@

Since the novels are chosen based on their worldwide reception and canon, practical constraints like how all novels should be of similar length are ignored. This constraint can prevent one novel having more words than others, which can critically affect the integrity of the word frequency analysis, sentiment value and word networks where one novel dominates the others and skews the results.

In addition, these authors work are expansive and so choosing a select few to bind them to the SOC genre can be limiting as the SOC genre in itself is already an amalgamation of different literary trends. Thus, this can affect the integrity of the datasets.

@@

Along the search for other stream of consciousness authors to include in the generated datasets, only a few-those that are in the SOC canon, mostly white authors-are present in Project Gutenberg, without including lesser known SOC authors, which could have steared the data analysis, still, towards their preconception: being rueful and angst-ridden
